Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-03 overleaf/related.tex
Fixed refs.

\paragraph{MPC languages and compilers.}
Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
as it falls between the high-level language and lower-level circuit protocols.

@@

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All (\url{https://github.com/Par4All/par4all}), however, does not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops. %and achieve significant speedup.

>>>>

\paragraph{MPC languages and compilers.}
Languages and compilers for secure computation have seen significant advances in recent years. The early MPC compilers Fairplay~\cite{Ben-David:2008},
and Sharemind~\cite{Bogdanov:2008} were followed by PICCO~\cite{Zhang:2013}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{Songhori:2015},
Wystiria~\cite{Rastogi:2014}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{Demmler:2015,Buscher:2018b,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
as it falls between the high-level language and lower-level circuit protocols.

@@

HyCC~\cite{Buscher:2018b} is a compiler from C Source into ABY circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All (\url{https://github.com/Par4All/par4all}), however, does not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- concretely, it is well-known that vectorization and polyhedral compilation 
do not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops. %and achieve significant speedup.

### 2022-05-03 overleaf/results.tex
Fixed refs.

We used benchmarks from ABY and OPA~\cite{NDSS:DemSchZoh15,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{CCS:BDKKS18} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{NDSS:DemSchZoh15} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{CCS:BDKKS18} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

@@

HyCC~\cite{CCS:BDKKS18} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

>>>>

We used benchmarks from ABY and OPA~\cite{Demmler:2015,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{Buscher:2018b} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{Demmler:2015} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{Buscher:2018} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

@@

HyCC~\cite{Buscher:2018b} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

### 2022-05-03 overleaf/vectorization.tex
Fixed refs.

This section describes our vectorization algorithm. While vectorization is a longstanding problem, 
and we build upon existing work on scalar expansion and classical loop vectorization~\cite{Allen:1987}, 
our algorithm is unique as it works on the MPC Source SSA-form representation. We posit that vectorization 
over MPC Source is a new problem that warrants a new look, in part because of MPC's unique linear 
structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).

>>>>

This section describes our vectorization algorithm. While vectorization is a longstanding problem, 
and we build upon existing work on scalar expansion and classical loop vectorization~\cite{Allen:1987}, 
our algorithm is unique as it works on the MPC Source SSA-form representation. We posit that vectorization 
over MPC Source is a new problem that warrants a new look, in part because of MPC's unique linear 
structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{Ishaq:2019}).
